---
title: "Code Book"
output: pdf_document
---
##Loading the Files and completing the data sets
Loading the Libraries that are going to be used:

```{r Reading Library, message=FALSE}
library(dplyr) #Use for manipulating the data
```

Reading the data:

```{r Reading the Xtest and Xtrain files, cache=TRUE}
Xtest  <- tbl_df(read.table(file = "test/X_test.txt"))
Xtrain <- tbl_df(read.table(file = "train/X_train.txt"))

Ytest <- read.table("test/y_test.txt")
Ytrain <- read.table("train/y_train.txt")
SubjectTest <- read.table("test/subject_test.txt")
SubjectTrain <- read.table("train/subject_train.txt")
Features <- read.table("features.txt")
ActLabel <- read.table("activity_labels.txt")

```

The code bellow show how the test and train data set are completed with other information from the folder.

```{r Manipulating Data}
colnames(Xtest) <- Features[,2]##Add the Features for Test
colnames(Xtrain)<- Features[,2]##Add the Features for Train

Xtest$Activity <- Ytest[,1]  ##Activity ID for Test
Xtrain$Activity <- Ytrain[,1] ##Activity ID for Train
Xtest$Subject <- SubjectTest[,1] ##Subject ID for Test
Xtrain$Subject <- SubjectTrain[,1] ##Subject ID for Train
```

##Assignment 1: Merging the data sets

For the first assignment we are supposed to merge the two data sets: Xtest and Xtrain  
If there are duplicated variables, they are removed.
 
```{r Merging the datasets, cache=TRUE}
merge <- tbl_df(rbind(Xtest,Xtrain))
if(sum(duplicated(colnames(merge)))>0){
        merge <- merge[,!duplicated(colnames(merge))] ##Remove duplicated Variables
}
```


##Assignment 2: Extract the variable of the mean and standart deviation
This chunk of code goes though the variable names and only select the ones that have mean or std in their names.  

```{r Mean and Standard Deviation}
Xbar <- merge[,grep("[Mm][e][a][n]",colnames(merge))]#Gets all the variables that have mean
Sbar <- merge[,grep("[Ss][t][d]",colnames(merge))]#Gets all the variables that have Std
```

##Assignment 3: Names the types of activities

Create an inside function in sapply to name and create factors for the Activity Variable

```{r Text variable, echo=FALSE}
merge$Activity <- sapply(X = merge$Activity, y = ActLabel,
                         function(x, y){x <- y[x,2]})
head(merge$Activity)
```

##Assignment 4: Give readable names for the variables 
By the file features_info, there is somea breviations  
f: Frequency; t:Time ; Mag: Magnitude; Acc: Accelerator; Gyro:Gyroscope

```{r Replancing old names for new ones}
names(merge) <- gsub("^f","Frequency",names(merge))
names(merge) <- gsub("^t","Time",names(merge))
names(merge) <- gsub("Mag","MAgnitude",names(merge))
names(merge) <- gsub("Acc","Accelerator",names(merge))
names(merge) <- gsub("Gyro","Gyroscope",names(merge))
```

The participants are identified with the following code   

```{r naming the participants}
merge$Subject <- as.factor(sapply(X = merge$Subject,
                           function(x){x <- paste("Participant", x, sep = " ")}))
head(merge$Subject)
```

##Assignment 5: Create a Tidy Data with the mean by Activity and Subject

The functions *group_by* and *summarize_each* from the dplry package are used to improve the process  

```{r Groupping and summarizing}
merge <- group_by(merge, Subject, Activity)
TidyData <- summarize_each(merge, funs(mean))
write.table(TidyData, file = "Tidy.txt", row.names = FALSE)
```

That is the process taken for tyding the data.

```{r Removing the variables created for manipulation, echo=FALSE}
a = NULL
a <- ls()
a <- a[which(a!="TidyData" & a!="Xtest" & a!="Xtrain")]
rm(list = a)
```